Estimating Learning Dynamics in the Routing Game

### 2015-10-14 paper/paper.tex

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\begin{figure*}
  \centering
  \includegraphics[width=160mm]{images/user_interface}
  \caption{Web interface for the routing game.}
  \label{fig:web_interface}
\end{figure*}
%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Setup}

We simulate the proposed methods with a routing game system that implements the behavior of the network. Players are assigned a unique origin and destination node pair and are tasked to assign a flow distribution to the paths between this pair at each turn of the game. The game calculates the cost along each edge of the network after all players have submitted their flow distribution for the turn. Each player then receives feedback with the local information of the cost for each path. The players then use this path cost history to assign a new flow distribution at the subsequent turns with the objective of minimizing their cumulative path cost.

We simulate the proposed methods with a routing game interface that implements the behavior of the network. Players are assigned a unique origin and destination node pair and are tasked to assign a flow distribution to the paths between this pair at each turn of the game. The game calculates the cost along each edge of the network after all players have submitted their flow distribution for the turn. Each player then receives feedback with the local information of the cost for each path. The players then use this path cost history to assign a new flow distribution at the subsequent turns with the objective of minimizing their cumulative path cost.

@@

We can estimate the learning rate of each player with the update in Figure~\ref{eq:estimation_eta}. We can see that the learning rates of some players are negative at certain turns, which may be explained by the players assuming a contrarian play style at flow distribution assignments for those certain turns.

\begin{figure}
  \centering
  \includegraphics[width=80mm]{images/learning_rates}
  \caption{Learning rate for each player.}
  \label{fig:learning_rates}
\end{figure}

With these estimated sequence of learning rates for each player, we can predict the flow distribution at turn $t+1$ with the learning rate $\eta_t$. We propose three methods for calculating $\eta_t$: First we let $\eta_t=\eta_{t-1}$, second we compute a moving average
\[
\eta_t=\dfrac{1}{N}\sum_{i=t-N-1}^{t-1}\eta_i
\] and finally we compute $\eta_t$ as a linear regression from $\eta_i,\ i \in [0,\ t-1]$.

>>>>

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
% \begin{figure*}
%   \centering
%   \includegraphics[width=160mm]{images/user_interface}
%   \caption{Web interface for the routing game.}
%   \label{fig:web_interface}
% \end{figure*}
%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Setup}

We simulate the proposed methods with a routing game system that implements the behavior of the network. The system is implemented with a web framework that is accessible to players through their browser. Players are assigned a unique origin and destination node pair and are tasked to assign a flow distribution to the paths between this pair at each turn of the game. The game calculates the cost along each edge of the network after all players have submitted their flow distribution for the turn. Each player then receives feedback with the local information of the cost for each path. The players then use this path cost history to assign a new flow distribution at the subsequent turns with the objective of minimizing their cumulative path cost.

\begin{figure}
  \centering
  %   \includegraphics[width=0.5\textwidth]{images/fill_me}
  \missingfigure[figwidth=6cm]{System block diagram}
  \caption{System block diagram to be added later.}
  \label{fig:system_block_diagram}
\end{figure}

% \begin{figure}
%   \centering
%   \includegraphics[width=0.5\textwidth]{images/game_system}
%   \caption{Player normalized cost.}
%   \label{fig:normalized_costs}
% \end{figure}

% \tikzstyle{block} = [draw, fill=white, rectangle,
% minimum height=3em, minimum width=6em]

% \tikzstyle{block} = [draw,fill=blue!20,minimum size=2em]

% \tikzstyle{sum} = [draw, fill=white, circle, node distance=1cm]
% \tikzstyle{input} = [coordinate]
% \tikzstyle{output} = [coordinate]
% \tikzstyle{pinstyle} = [pin edge={to-,thin,black}]

% \def\radius{.7mm}
% \tikzstyle{branch}=[fill,shape=circle,minimum size=3pt,inner sep=0pt]

% \begin{tikzpicture}[auto, node distance=2cm,>=latex']

%   \node [input, name=input] {};
%   \node [sum, right of=input] (sum) {};
%   % \node [block, right of=sum] (controller) {Controller};
%   \node [block, right of=sum, node distance=3cm] (system) {Network};

%   \node [output, right of=system] (output) {};
%   % \node [block, below of=system] (measurements) {Measurements};

%   % \draw [draw,->] (input) -- node {$\bar x(t)$} (sum);
%   \draw [->] (sum) -- node {$\bar x(t)$} (system);
%   % \draw [->] (y) |- (measurements);
%   % \draw [->] (measurements) -| node[pos=0.99] {$-$}
%   % node [near end] {$y_m$} (sum);

%   \node at (0,-1) (input1) {$i_1$};
%   \node[block, above left = of system, anchor=west] at (2,-1) (block1) {$p_1$};
%   \draw[->] (block1.east) -- (sum);
%   \draw[->] (system) -- (block1);

%   \node at (0,-2) (input2) {$i_2$};
%   \node[block, left of=system, anchor=west] at (2,-2) (block2) {$p_2$};
%   \draw[->] (block2.east) -- (sum);
%   \draw[->] (system) -- (block2);

%   \node at (0,-3) (input3) {$i_3$};
%   \node[block, below left = of system, anchor=west] at (2,-3) (block3) {$p_3$};
%   \draw[->] (block3.east) -- (sum);
%   \draw[->] (system) -- (block3);

%   % % Draw blocks, inputs and outputs
%   % \foreach \y in {1,2,3} {
%   %   \node at (0,-\y) (input\y) {$i_\y$};
%   %   \node[block, left of=system, above of =system] at (2,-\y) (block\y) {$p_\y$};
%   %   \draw[->] (input\y) -- (block\y);
%   %   \draw[->] (block\y.east) -- (sum);
%   %   \draw[->] (system) -- (block\y);
%   % }

%   % Calculate branch point coordinate
%   \path (input1) -- coordinate (branch) (block1);

%   % Define a style for shifting a coordinate upwards
%   % Note the curly brackets around the coordinate.
%   \tikzstyle{s}=[shift={(0mm,\radius)}]
%   % It would be natural to use the yshift or xshift option, but that does
%   % not seem to work when shifting coordinates.

%   \draw[->] (branch) node[branch] {}{ % draw branch junction
%     \foreach \c in {2} {
%       % Draw semicircle junction to indicate that the lines are
%       % not connected. The intersection between the lines are
%       % calculated using the convenient -| syntax. Since we want
%       % the semicircle to have its center where the lines intersect,
%       % we have to shift the intersection coordinate using the 's'
%       % style to account for this.
%       [shift only] -- ([s]input\c -| branch) arc(90:-90:\radius)
%       % Note the use of the [shift only] option. It is not necessary,
%       % but I have used it to ensure that the semicircles have the
%       % same size regardless of scaling.
%     }
%   };

% \end{tikzpicture}

To address a player who may be at a disadvantage due to having longer paths or more congested paths because of the topology of the network, we normalized the cumulative path cost of each player with their cumulative path cost at equilibrium calculated with Algorithm~\ref{alg:MD}.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{images/players_costs}
  \caption{Player cost normalized by equilibrium cost .}
  \label{fig:normalized_costs}
\end{figure}

\section{Analysis}
\label{sec:analysis}

@@

We can estimate the learning rate of each player with the update in Eq~\ref{eq:estimation_eta}. We obtained these sequence of learning rates for each player and for better visualization, we plot the moving average

% \begin{figure}
%   \centering
%   \includegraphics[width=0.5\textwidth]{images/learning_rates}
%   \caption{Learning rate for each player.}
%   \label{fig:learning_rates}
% \end{figure}

With these estimated sequence of learning rates for each player, we can predict the flow distribution at turn $t+1$ with the learning rate $\eta_t$.
% \[
% \eta_t=\dfrac{1}{N}\sum_{i=t-N-1}^{t-1}\eta_i
% \] and finally we compute $\eta_t$ as a linear regression from $\eta_i,\ i \in [0,\ t-1]$.
